Modular Curves

### 2020-01-21 paper.tex
lots of stuff added

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
%\usepackage[dutch]{babel}
\usepackage{amsmath, amssymb, amsthm}
%\usepackage{geometry}
\usepackage{tikz-cd}
\usepackage{mdframed}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{todonotes}
% figure support
\usepackage{import}
\usepackage{xifthen}
\usepackage{pdfpages}
\usepackage{transparent}
\newcommand{\incfig}[1]{%
	\def\svgwidth{\columnwidth}
	\import{./figures/}{#1.pdf_tex}
}

@@

\theoremstyle{theoremdd}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{property}[theorem]{Property}
\theoremstyle{definitiondd}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remarkdd}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\ltr}{\par \noindent \framebox[1\width]{ $\implies$ } \hspace{.2cm}}
\newcommand{\rtl}{\par \noindent \framebox[1\width]{ $\impliedby$ } \hspace{.2cm} }
\newcommand{\exercise}[1]{\begin{mdframed}
	\textit{Exercise:} #1
\end{mdframed}}
\newcommand{\bigset}[2]{ \left\{ #1 \;\middle|\; #2 \right\} }

@@

This is the report of the journey of a student learning about elliptic curves. 
This report should be thought of as my notes while working my way through the first chapters of \emph{A First Course in Modular Forms} by Fred Diamond and Jerry Shurman. 
This book is my main reference. 
Most of this report is paraphrasing this book, but I've added some details where I thought the book was lacking. 
I've also tried to add some visual intuitions I developed to aid my understanding of this material. I hope potential readers will find these useful as well.

@@

\subsection{Subgroups of the Modular Group}
Quite often it will be useful to restrict our focus to subgroups of the modular group. 
The type of subgroup we'll most interested in are the so-called \emph{congruence subgroups}.
To define these we first have to define the so called \emph{principal congruence subgroups}.  
\begin{definition}
	For any $N \in \N$ we define the \emph{principal congruence subgroup of level $N$} to be \[
		\Gamma(N) = \bigset{\begin{pmatrix} a & b \\ c & d \end{pmatrix} \in \SL_2(\Z)}{\begin{pmatrix} a & b \\ c & d \end{pmatrix}  \equiv \begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix}  (\mathrm{mod}\; N)}
	.\] 
\end{definition}

@@

\section{Lattices and Complex Elliptic Curves}
The main idea of this section is to construct tori by considering $\C$ as an additive group modulo a certain kind of embedding of $\Z^2$. This embedding of $\Z^2$ is called a lattice.  
\begin{definition}
	A \emph{lattice in $\C$} is additive subgroup $\Lambda = \omega_1 \Z \oplus \omega_2 \Z$ where $\omega_1, \omega_2 \in \C$ are linearly independent over $\R$ and $\omega_1 / \omega 2  \in \mathbb{H}$
\end{definition}
\begin{definition}
	A \emph{complex torus} is a quotient of the complex plane by a lattice, \[
		\C / \Lambda = \bigset{z + \Lambda}{z \in \C} 
	.\] 
\end{definition}
This object is called as torus because of a geometric intuition. 
If one considers the parallelogram spanned by $\vec{\omega_1}, \vec{\omega_2}$ we see that the opposing edges are `glued' together. 
So topologically this space is homeomorphic to the torus.

\begin{figure}[h]
	\centering
	\incfig{lattice_complex_plane}
\end{figure}

@@

\begin{example}
	The so called \emph{multiply by integer maps} are isogenies. These are maps of the form \[
		[N]: \C / \Lambda \to \C / \Lambda: z + \Lambda \mapsto  Nz + \Lambda
	,\] where $N$ is an integer. 
	Note that the kernel of this map is $\frac{1}{N}\Lambda / \Lambda$ and is isomorphic to $\Z / N \Z \times  \Z / N \Z$. We denote this kernel as $E[N]$ and is the set of all $N$ torsion points in $\C / \Lambda$. This can be seen in \cref{fig:Ntorsion_subgroup}.
\end{example}
	\begin{figure}[h]
		\incfig{NTorsion_subgroup}	
		\caption{The $N$-torsion subgroup $E[N]$ of a complex torus, generated by  $\omega_1 / N$ and $\omega_2 / N$.}
		\label{fig:Ntorsion_subgroup}
	\end{figure}
\begin{example}
	Another example are the \emph{Cyclic quotient maps}. 
	Let  $\C / \Lambda$ be a complex torus and  $N \in \N$.
	If take an $N$ torsion point, $\lambda$, of $\C / \Lambda$ (a point in $\frac{1}{N} \Lambda / \Lambda$) then we can consider the subgroup $C$ generated by $\lambda$. Note that $C + \Lambda$ is a super lattice of $\Lambda$. Then the map \[
		\pi: \C \ \Lambda \to \C / (C + \lambda): z + \Lambda \to z + C + \Lambda
	\]  
	is an isogeny with kernel $C$. 
\end{example}

These two examples are very important, as they allow the construction of every isogeny. 
\begin{theorem}
	Let $\phi: \C / \Lambda \to \C / \Lambda': z + \Lambda \mapsto mz + \Lambda'$ be an isogeny. Then $\phi = \psi \circ \pi \circ [n]$, where  $\psi$ is an isomorphism, $\pi$ is a cyclic quotient map and $[n]$ is a multiply by integer map.
\end{theorem}
\begin{proof}
	Denote $K = \ker \phi$. We know that $K = m^{-1} \Lambda' / \Lambda$.
	Let $N$ be the order of $K$. So we know that every point of $K$ is $N$ torsion. 
	Hence \[
		K \subset N^{-1} \Lambda / \Lambda \simeq \Z / N \Z \oplus \Z / N \Z. 
	\]
	By the structure theorem for finite abelian groups we know that  \[
	K \cong \Z / n \Z \times \Z / n n' \Z
	\]
	for some $n, n'\in \N$. 
	This means that $nK \cong \Z / n' \Z$. Denote $\pi: \C / \Lambda \to \C / (nK + \Lambda)$ the cyclic quotient map. Now define the map \[
		\psi: \frac{\C}{nK + \Lambda} \to \frac{\C}{\Lambda'}: z + nK \mapsto  \left(\frac{m}{n}\right) z + \left( \frac{m}{n} \right)(nK + \lambda)
	,\] 
	which is a well defined isomorphism as $mK + m\Lambda = \Lambda'$. 
	I now claim that $\phi = \psi \circ \pi \circ [n]$. Indeed, take any $z + \Lambda$ then \[
		z + \Lambda \mapsto nz + \Lambda \mapsto  nz + nK + \Lambda \mapsto  mz + \Lambda' \phi(z + \Lambda)
	.\]  
\end{proof}
\todo{dual isogenies, weil pairings}
\section{Ellpitic curves and their connection to Complex Tori}
An elliptic curve is a cubic curve of the form \[
y^2 = 4x^3 - g_2x - g_3
,\] 
where $g_2, g_3 \in \Z$ and $g_2^3 - 27 g_3^2 \ne 0$. 
These objects are related to complex tori, by the meromorphic functions from the complex tori to the complex plane. 
A mermorphic function $\phi: \C / \Lambda \to \C \cup \{\infty\} $ are naturally identified with a  $\Lambda$-periodic function $\phi: \C \to \C \cup \{\infty\}$.
One may wonder how a non-trivial example of such a function looks like. 
One example of particular interest is the \emph{Weierstrass $\wp$-function}. 
For a given lattice  $\Lambda$ we define this function to be \[
	\wp(z) = \frac{1}{z^2} + \sum_{\substack{\omega \in \Lambda \\ \omega \ne 0}}\left( \frac{1}{(z-\omega)^2} - \frac{1}{\omega^2} \right) 
.\] 
At first glance it seems obvious that such a function is $\Lambda$ periodic, but we have to be careful. 
While this sum is absolutely convergent on compact sets, its is not if we split up $\frac{1}{(z-\omega^2)} - \frac{1}{\omega^2}$ into two different summations. So we cannot simply permute the terms in the desired way. 
It still is $\Lambda$-periodic. Indeed, notice that  \[
	\wp'(z) = -2 \sum_{\omega \in \Lambda} \frac{1}{(z - \omega)^3}
\] 
this is periodic function as we can permute the terms in the sum.
We will now show that this implies that $\wp$ is $\Lambda$-periodic (exercise 1.4.2).
Note that $\wp$ is symmetric, as showing this can indeed be achieved by permuting the right terms.
For any $\omega \in \Lambda$ is holds that 
\[
	(\wp(z) - \wp(z + \omega))' = \wp'(z) - \wp'(z+ \omega) = 0 
.\] 
From this it follows that $\wp(z) - \wp(z + \omega)$ is constant a constant function of $z$, everywhere where it is defined.
In particular it is the case that \[
	\wp\left(-\frac{\omega}{2}\right) - \wp\left(\omega - \frac{\omega}{2}\right) = \wp\left(\frac{\omega}{2}\right) - \wp\left(\frac{\omega}{2}\right)  = 0 
.\] 
So $\wp(z) = \wp(z+ \omega)$. 
We conclude that $\wp$ is $\Lambda$-periodic.

@@

Recall that \emph{Eisenstein-series} $G_k$ is the function defined as \[
	G_k(\tau) = \sum_{\substack{c, d \in \Z^2 \\ (c, d) \ne 0}} \frac{1}{(c\tau + d)^{k}}
.\] 
Note that this is equal to \[
	G_k(\tau) = \sum_{\substack{\omega \in \Lambda_\tau \\ \omega \ne 0}} \frac{1}{\omega}
.\] 
In this spirit we extend the definition of the eisentein series as follows. For a given lattice $\Lambda$ and an even number  $k > 2$ we define \[
	G_k(\Lambda) = \sum_{\substack{\omega \in \Lambda \\ \omega \ne 0}} \frac{1}{\omega^{k}}
.\] 
Note that this is a very natural extension as $G_k(\tau) = G_k(\Lambda_\tau)$.
\section{Appendices}
\begin{theorem}
	The modular group $\SL_2(\Z)$ is generated by \[ 	
	\begin{pmatrix}  1 & 1 \\ 0 & 1 \end{pmatrix} 
	\text{ and }
	\begin{pmatrix} 0 & -1 \\ 1 & 0 \end{pmatrix} 
.\] 
\end{theorem}
\begin{proof}
	Let $\mathcal{A} $ be the group generated by the two matrices. 
	Suppose that $\SL_2(\Z) \setminus \mathcal{A} $ is not empty.
	Then we can choose a matrix \[
		A =\begin{pmatrix} a & b \\ c& d\end{pmatrix} \in \SL_2\Z \setminus \mathcal{A} 
	\] 
	with mimimal 1-norm, i.e. $|a| + |b| + |c| + |d|$ is minimal.
	We know that 
\end{proof}
\end{document}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{textcomp}
%\usepackage[dutch]{babel}
\usepackage{amsmath, amssymb, amsthm}
%\usepackage{geometry}
\usepackage{tikz-cd}
\usepackage{mdframed}
\usepackage{microtype}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{todonotes}
\usepackage[backend=biber]{biblatex}
\addbibresource{references.bib}
% figure support
\usepackage{import}
\usepackage{xifthen}
\usepackage{pdfpages}
\usepackage{transparent}
\newcommand{\incfig}[1]{%
	\def\svgwidth{\columnwidth}
	\import{./figures/}{#1.pdf_tex}
}

@@

\theoremstyle{theoremdd}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{property}[theorem]{Property}
\theoremstyle{definitiondd}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remarkdd}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{exercise}[theorem]{Exercise}

\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\ltr}{\par \noindent \framebox[1\width]{ $\implies$ } \hspace{.2cm}}
\newcommand{\rtl}{\par \noindent \framebox[1\width]{ $\impliedby$ } \hspace{.2cm} }
\newcommand{\bigset}[2]{ \left\{ #1 \;\middle|\; #2 \right\} }

@@

This is the report of the journey of a student learning about elliptic curves. 
This report should be thought of as my notes while working my way through the first chapters of \emph{A First Course in Modular Forms} by Fred Diamond and Jerry Shurman\cite{diamondFirstCourseModular2005a}. 
This book is my main reference. 
Most of this report is paraphrasing this book, but I've added some details where I thought the book was lacking. 
I've also tried to add some visual intuitions I developed to aid my understanding of this material. I hope potential readers will find these useful as well.

@@

\subsection{Subgroups of the Modular Group} \label{sec:subgroups_of_the_modular_group}
Quite often it will be useful to restrict our focus to subgroups of the modular group. 
The type of subgroup we'll most interested in are the so-called \emph{congruence subgroups}.
To define these we first have to define the so called \emph{principal congruence subgroups}.  
\begin{definition}
	For any $N \in \N$ we define the \emph{principal congruence subgroup of level $N$} to be \[
		\Gamma(N) = \bigset{\begin{pmatrix} a & b \\ c & d \end{pmatrix} \in \SL_2(\Z)}{\begin{pmatrix} a & b \\ c & d \end{pmatrix}  \equiv \begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix}  (\mathrm{mod}\; N)}
	.\] 
\end{definition}

@@

\section{Lattices and Complex Elliptic Curves}
The main idea of this section is to construct tori by considering $\C$ as an additive group modulo a certain kind of embedding of $\Z^2$. This embedding of $\Z^2$ is called a lattice.  
\begin{definition}
	A \emph{lattice in $\C$} is additive subgroup $\Lambda = \omega_1 \Z \oplus \omega_2 \Z$ where $\omega_1, \omega_2 \in \C$ are linearly independent over $\R$ and $\omega_1 / \omega 2  \in \mathbb{H}$
\end{definition}
\begin{definition}
	A \emph{complex torus} is a quotient of the complex plane by a lattice, \[
		\C / \Lambda = \bigset{z + \Lambda}{z \in \C} 
	.\] 
\end{definition}
This object is called as torus because of a geometric intuition. 
If one considers the parallelogram spanned by $\vec{\omega_1}, \vec{\omega_2}$ we see that the opposing edges are `glued' together. 
So topologically this space is homeomorphic to the torus (see \cref{fig:complex_torus}).

\begin{figure}[h]
	\centering
	\incfig{lattice_complex_plane}
	\caption{A lattice generated by $\omega_1$ and $\omega_2$. The grey part shows a fundamental parallelogram of the lattice which represents the complex torus $\C / \Lambda$. The arrows on the edges shows how the edges are `glued' together to form a torus.} \label{fig:complex_torus}
\end{figure}

@@

\begin{example}
	The so called \emph{multiply by integer maps} are isogenies. These are maps of the form \[
		[N]: \C / \Lambda \to \C / \Lambda: z + \Lambda \mapsto  Nz + \Lambda
	,\] where $N$ is an integer. 
	Note that the kernel of this map is $\frac{1}{N}\Lambda / \Lambda$ and is isomorphic to $\Z / N \Z \times  \Z / N \Z$. We denote this kernel as $E[N]$ and is the set of all $N$ torsion points in $\C / \Lambda$. This can be seen in \cref{fig:Ntorsion_subgroup}.
\end{example}
	\begin{figure}[h]
		\incfig{NTorsion_subgroup}	
		\caption{The $N$-torsion ($N = 4$) subgroup $E[N]$ of a complex torus, generated by  $\omega_1 / N$ and $\omega_2 / N$.}
		\label{fig:Ntorsion_subgroup}
	\end{figure}
\begin{example}
	Another example are the \emph{Cyclic quotient maps}. 
	Let  $\C / \Lambda$ be a complex torus and  $N \in \N$.
	If take an $N$ torsion point, $\lambda$, of $\C / \Lambda$ (a point in $\frac{1}{N} \Lambda / \Lambda$) then we can consider the subgroup $C$ generated by $\lambda$. Note that $C + \Lambda$ is a super lattice of $\Lambda$. Then the map \[
		\pi: \C \ \Lambda \to \C / (C + \lambda): z + \Lambda \to z + C + \Lambda
	\]  
	is an isogeny with kernel $C$. 
\end{example}

These two examples are very important, as they allow the construction of every isogeny. 
\begin{theorem}
	Let $\phi: \C / \Lambda \to \C / \Lambda': z + \Lambda \mapsto mz + \Lambda'$ be an isogeny. Then $\phi = \psi \circ \pi \circ [n]$, where  $\psi$ is an isomorphism, $\pi$ is a cyclic quotient map and $[n]$ is a multiply by integer map.
\end{theorem}
\begin{proof}
	Denote $K = \ker \phi$. We know that $K = m^{-1} \Lambda' / \Lambda$.
	Let $N$ be the order of $K$. So we know that every point of $K$ is $N$ torsion. 
	Hence \[
		K \subset N^{-1} \Lambda / \Lambda \simeq \Z / N \Z \oplus \Z / N \Z. 
	\]
	By the structure theorem for finite abelian groups we know that  \[
	K \cong \Z / n \Z \times \Z / n n' \Z
	\]
	for some $n, n'\in \N$. 
	This means that $nK \cong \Z / n' \Z$. Denote $\pi: \C / \Lambda \to \C / (nK + \Lambda)$ the cyclic quotient map. Now define the map \[
		\psi: \frac{\C}{nK + \Lambda} \to \frac{\C}{\Lambda'}: z + nK \mapsto  \left(\frac{m}{n}\right) z + \left( \frac{m}{n} \right)(nK + \lambda)
	,\] 
	which is a well defined isomorphism as $mK + m\Lambda = \Lambda'$. 
	I now claim that $\phi = \psi \circ \pi \circ [n]$. Indeed, take any $z + \Lambda$ then \[
		z + \Lambda \mapsto nz + \Lambda \mapsto  nz + nK + \Lambda \mapsto  mz + \Lambda' \phi(z + \Lambda)
	.\]  
\end{proof}

\todo{maybe say something about dual maps}
\subsection{Weil Pairings} \label{sec:Weil_Pairings}
Let $E[N]$ be the  $N$ torsion subgroup of some complex torus $\C / \Lambda$, with $\Lambda = \omega_1\Z + \omega_2 \Z$. 
The \emph{Weil pairing} is a map \[
	e_N: E[N] \times E[N] \to \Z / N\Z
.\] 
Although this object comes up a lot in de discussion of complex tori/complex elliptic curves, it is defined very algebraically. 
Recall that $E[N]$ is generated by $\omega_1 / N$ and $\omega_2 / N$. So every point in $E[N]$ is  $a \cdot \omega_1 / N + b \cdot \omega_2 / N$ for some integers $a, b$ which are defined up to modulo $N$. In essense we may say that $(a, b)$ are the coordinates of point in $E[N]$. 
Now the map Weil pairing is defined as the map
\begin{align*}
	e_n: E[N] \times  E[N] &\longrightarrow \frac{\Z}{N\Z} \\
	\left(a \frac{\omega_1}{N} + b \frac{\omega_2}{N}, c \frac{\omega_1}{N} + d \frac{\omega_2}{N}\right) &\longmapsto ac-bd =  \det \begin{pmatrix} a & b \\ c& d \end{pmatrix} 
.\end{align*}
\begin{figure}[h]
	\incfig{weilpairing}
	\caption{A fun geometric interpretation of the weilpairing}
	\label{fig:weilpairing}
\end{figure}
A fun geometric interpretation of the weil pairing is given in \cref{fig:weilpairing}. As the Weil pairing is a determinant there is a geometric interpretation of it as an area. In light of Picks theorem \cite{wiki:pick} we see that the Weial pairing is the number of points in the shaded region in the figure modulo $N$. 
From this interpretation it is imediately obvious that the Weil pairing is independent of the choice of basis for  $\Lambda$ as long as the orientation condition $\frac{\omega_1}{\omega_2} \in \mathcal{H}$ is satisfied. 
From the definition it is clear that $e_N$ is bilinear and skew symmetric.
As the Weil pairing is completely defined on the  $\Z / N \Z \times  \Z / N\Z$ group structure of $E[N]$ it is preserved under isomorphism.
\section{Ellpitic curves and their connection to Complex Tori}
\subsection{What are Elliptic Curves}
This section is based on section 1.2 of the book \emph{Rational Points on Elliptic Curves} by Joseph H.\ Silverman and John T.\ Tate\cite{silvermanRationalPointsElliptic2015}.
An elliptic curve is a cubic curve of the form \begin{equation}\label{eq:elliptic_curve}
y^2 = 4x^3 - g_2x - g_3
,\end{equation}
where $g_2, g_3 \in \Z$ and $g_2^3 - 27 g_3^2 \ne 0$. 
While we do not write this as an homogeneous equation we usually consider this to be its projective closure in $\mathbb{P}^2$. 
One of the most important properties of elliptic curves is that they have a certain natural abelian group structure. 
We first have have to choose a zero point, $\mathcal{O} $, usually we choose the point at infinity.
Recall B\'ezout's theorem. 
This states that any line intersects the curve \eqref{eq:elliptic_curve} in exactly 3 points, counting multiplicity.  
So for any two points $P$ and $Q$ on the curve there is exactly one line, $\ell$ that that intersects the curve in both $P$ and $Q$ (the intersection number is 2  if $P = Q$). 
Then there is a unique third intersection of this points. Let we denote this intersection by $P*Q$ be this intersection.
We can again take the third intersection of the the line through $P*Q$ and $\mathcal{O}$. This intersection is the result of $P + B$.
In essence we define define $P + Q =\mathcal{O} *(P*Q)$.
In the case where $\mathcal{O} $ is the point at infinity, the operation $\mathcal{O} * -$ is mirroring about the $x$ axis. So if $P * Q = (x,y)$, then $P + Q = (x, -y)$.
It can definitely be shown directly that this is a well defined abelian group structure. But we will show this in a more indirect way.

In the next section we will bijects the points on the curve \eqref{eq:elliptic_curve} with the points of a certain complex torus and show that the group action of the elliptic curve corresponds with the addition on the complex torus.

\subsection{The Weierstrass $\wp$ Function} \label{sec:The_Weierstrass_wp_Function}
We now continue our discussion of \cite{diamondFirstCourseModular2005a}.

Elliptic curves are related to complex tori, by the meromorphic functions from the complex tori to the complex plane. These functions are also called \emph{elliptic functions}. 
A elliptic function $\phi: \C / \Lambda \to \C \cup \{\infty\} $ is naturally identified with a  $\Lambda$-periodic function $\phi': \C \to \C \cup \{\infty\}: z \mapsto \phi(z + \Lambda)$. 
We will often make use of this identification implicitly. 
One may wonder how a non-trivial example of such a function looks like. 
One example of particular interest is the \emph{Weierstrass $\wp$-function}. 
For a given lattice  $\Lambda$ we define this function to be \[
	\wp(z) = \frac{1}{z^2} + \sum_{\substack{\omega \in \Lambda \\ \omega \ne 0}}\left( \frac{1}{(z-\omega)^2} - \frac{1}{\omega^2} \right) 
.\] 
At first glance it seems obvious that such a function is $\Lambda$ periodic, but we have to be careful. 
While this sum is absolutely convergent on compact sets, its is not if we split up $\frac{1}{(z-\omega^2)} - \frac{1}{\omega^2}$ into two different summations. So we cannot simply permute the terms in the desired way. 
It still is $\Lambda$-periodic. Indeed, notice that  \[
	\wp'(z) = -2 \sum_{\omega \in \Lambda} \frac{1}{(z - \omega)^3}
\] 
this is periodic function as we can permute the terms in the sum.
We will now show that this implies that $\wp$ is $\Lambda$-periodic (exercise 1.4.2).
Note that $\wp$ is symmetric, as showing this can indeed be achieved by permuting the right terms.
For any $\omega \in \Lambda$ is holds that 
\[
	(\wp(z) - \wp(z + \omega))' = \wp'(z) - \wp'(z+ \omega) = 0 
.\] 
From this it follows that $\wp(z) - \wp(z + \omega)$ is constant a constant function of $z$, everywhere where it is defined.
In particular it is the case that \[
	\wp\left(-\frac{\omega}{2}\right) - \wp\left(\omega - \frac{\omega}{2}\right) = \wp\left(\frac{\omega}{2}\right) - \wp\left(\frac{\omega}{2}\right)  = 0 
.\] 
So $\wp(z) = \wp(z+ \omega)$. 
We conclude that $\wp$ is $\Lambda$-periodic.

@@

Recall that \emph{Eisenstein-series} $G_k$ is the function defined as \[
	G_k(\tau) = \sum_{\substack{c, d \in \Z^2 \\ (c, d) \ne 0}} \frac{1}{(c\tau + d)^{k}}
.\] 
Note that this is equal to \[
	G_k(\tau) = \sum_{\substack{\omega \in \Lambda_\tau \\ \omega \ne 0}} \frac{1}{\omega}
.\] 
In this spirit we extend the definition of the eisentein series as follows. For a given lattice $\Lambda$ and an even number  $k > 2$ we define \[
	G_k(\Lambda) = \sum_{\substack{\omega \in \Lambda \\ \omega \ne 0}} \frac{1}{\omega^{k}}
.\] 
This is indeed a very natural generalisation as $G_k(\tau) = G_k(\Lambda_\tau)$.

We can describe the Laurent expansion of $\wp$ using these Eisenstein series.
\begin{proposition}\label{prop:laurent_weierstrass}
  Let $\Lambda$ be a lattice. Then the Laurent expansion of $\wp_\Lambda$ is \[
	  \wp_\Lambda(z) = \frac{1}{z^2} \sum_{\substack{n = 2 \\ n \text{ even}}}^{\infty}(n+ 1) G_{n + 2}(\Lambda)z^{n}
  .\] 
\end{proposition}
\begin{proof}
	As we are dealing with analytical functions it is sufficient to show equality on an non empty open. We choose a neighbourhood of 0 such that in that neighbourhood $|z| < |\omega|$ for all $\omega \in \Lambda \setminus \{ 0\} $. It that case it holds that. 
\begin{align*}
	\frac{1}{(z- \omega)^2} - \frac{1}{\omega^2} &= \frac{1}{\omega^2}\left( \left( \frac{1}{(1-z / \omega)} \right) ^2 - 1 \right)  \\
						     &= \frac{1}{\omega^2}\left( \left( \sum_{n = 0}^{\infty} \left(\frac{z}{\omega}\right)^{n} \right)^2 - 1  \right) \\
						     &= \frac{1}{\omega^2}\left( \sum_{n = 0}^{\infty} (n + 1)  \left( \frac{z}{\omega} \right) ^{n} - \right)  \\
						     &= \sum_{n = 1}^{\infty} (n + 1)  \frac{z^{n}}{\omega^{n+2}} \\
\end{align*}	
We can fill this in in the expression for $\wp_\Lambda$ yields
\begin{align*}
	\wp(z) &= 
	\frac{1}{z^2} + \sum_{\substack{\omega \in \Lambda \\ \omega \ne 0 }} \frac{1}{(z - \omega)^2} - \frac{1}{\omega^2} \\
	       &=  \frac{1}{z^2} + \sum_{\substack{\omega \in \Lambda \\ \omega \ne 0}} \sum _{n = 1}^{\infty} (n + 1)  \frac{z^{n}}{\omega^{n+2}}\\
	       &= \frac{1}{z^2} + \sum_{n = 1}^{\infty}\sum_{\substack{\omega \in \Lambda \\ \omega \ne 0 }} (n + 1) \frac{z^{n}}{\omega^{n+2}} \\
\end{align*}
We can exchange the order of summation as it turns out that this is an absolutely convergent series. 
When $n$ is odd we see that every term $ (n + 1) \frac{z^{n}}{\omega^{n+2}}$ cancels against $(n + 1) \frac{z^{n}}{(-\omega)^{n+2}}$. 
So we find that 
\begin{align*}
	\wp(z) &= \frac{1}{z^2} + \sum_{\substack{n = 2\\ n \text{ even}}}^{\infty}\sum_{\substack{\omega \in \Lambda \\ \omega \ne 0 }} (n + 1) \frac{z^{n}}{\omega^{n+2}} \\
	       &=\frac{1}{z^2} + \sum_{\substack{n = 2 \\ n \text{ even}}}^{\infty} (n+1) G_{n + 2}(\Lambda) z^{n}
.\end{align*}
\end{proof}
The function $\wp$ what connects complex tori to elliptic curves. 
\begin{theorem}\label{thm:complex_tori_elliptic_curves}
	Let $\Lambda = \omega_1 \oplus \omega_2$ be a lattice. Then for every $z + \Lambda$, the function  $(\wp_\Lambda(z), \wp'_\Lambda(z))$ is a point on the following cubic equation \begin{equation} \label{eq:cubic_curve_lattice}
		y^2 = 4x^3 - g_2(\Lambda)x - g_{3}(\Lambda)
	,\end{equation}
	where $g_2(\Lambda) = 60 G_4(\Lambda), g_3(\Lambda) = 140 G_6(\Lambda)$.
	The right hand side of this equation also factors in the following way \[
		y^2 = 4x^3 - g_2(\Lambda) x^2 - g_3(\Lambda) = 4\left(x - \wp_\Lambda\left(\frac{\omega_1}{2}\right)\right)\left(x - \wp_{\Lambda}\left(\frac{\omega_2}{2}\right)\right)\left(x - \wp\left(\frac{\omega_1 + \omega_2}{2}\right)\right)
	.\] 
\end{theorem}
\begin{proof}
	To prove \cref{eq:cubic_curve_lattice} we will first show that the equality holds for the first few terms of the Laurent expansion and then use a little trick using the periodicity of $\wp_\Lambda$ and $\wp'_\Lambda$.
	We know that (\cref{prop:laurent_weierstrass})\[
		\wp_\Lambda(z) = \frac{1}{z^2} + 3 G_4(\Lambda) z^2 + 5G_6(\Lambda) z^{4} + \mathcal{O} (z^{6})
	.\] 
	Similarly we see that \[
		\wp'_\Lambda = - \frac{2}{z^3} + 6 G_4(\Lambda)z + 20G_6(\Lambda)z^3 + \mathcal{O} (z^{5})	
	.\] 
	Using these expansions one can easily verify that \[
		(\wp'_\Lambda(z))^2 - 4(\wp(z))^3 + g_2(\Lambda)\wp_{\lambda}(z) + g_{3}(\Lambda)    = \mathcal{O} (z^{2})
	.\] 
	This expression is holomorphic and $\Lambda$-periodic. So it's bounded. We know that it must be zero as it is zero when  $z \to 0$. 
	As a result we find \cref{eq:cubic_curve_lattice}.

	We will now prove that the right-hand side factors in the desired way.
	As we've stated before $\wp$ is an even function. So $\wp'$ is an odd function. 
	If $z$ is in $E[2]$ then  $z + \Lambda = - z + \Lambda$. 
	Putting these two results together shows that $\wp'(z) = - \wp'(-z) = -\wp'(z)$.  
	As a result $\wp'(z) = 0$.
	Then we see that $\wp(z)$ must be a root of \cref{eq:cubic_curve_lattice}. 
	In particular  $\frac{\omega_1}{2}, \frac{\omega_2}{2}, \frac{\omega_1 + \omega_2}{2}$ are roots. 
	Because the derivatives $\wp'(\omega_1 / 2 ), \wp'(\omega_2 / 2), \wp'(\omega_1 / 2 + \omega_2 /2)$ are all $0$, \cref{ex:1.4.1} in the appendices shows that $\wp(\omega_1 / 2), \wp(\omega_2 / 2), \wp(\omega_1 / 2 + \omega_2 / 2)$ are distinct.
\end{proof}

Let's try to unpack the statement of \cref{thm:complex_tori_elliptic_curves}.
The first part shows that there is a map from a comples torus to an elliptic curve 
\begin{align*}
	\frac{\C}{\Lambda} &\longrightarrow \text{elliptic curve: } Z(x^3 - a x - b -y ^2) \\
	z+\Lambda &\longmapsto \left(\wp_\Lambda(z), \wp'_\Lambda(z)\right)
.\end{align*}
We know that this is an elliptic curve as it the right hand side has three distinct roots so its determinant $a^2 - 27 b^3 \ne 0  $.
We can see that this map is bijective. By \cref{ex:1.4.1}  $\wp_\Lambda$  takes every value twice. As this is an even function a certain $x \in \C$ must be taken twice, once by $z$ once by $-z$ and because $\wp'$ is an odd function $\wp'(z) \ne \wp'(-z)$ unless $z = 0$.  By for every $x$-value there are exactly two $y$ values on the elliptic curve. So this is indeed a bijection.

We will now show that this map is in fact a group isomorphism.
Let $z_1 + \Lambda, z_2 + \Lambda$ be points in $\C / \Lambda$.
These correspond to two points $P = (\wp_{\Lambda}(z_1), \wp'_\Lambda(z_1))$ and $ Q = (\wp_{\Lambda}(z_2), \wp'_\Lambda(z_2))$ in $\C^2$. 
Let $ax + by + c = 0$ be the line that intersects the elliptic curve in both these points (with intersection number 2 if  $z_1 = z_2$).
Define the function \[
	f(z) = a \wp_{\Lambda}(z) + b \wp'_\Lambda(z) + c
.\] 
Clearly $z_1, z_2$ are two zeros of this function. Obviously the third interesction of the line with the curve corresponds to the third zero of $f$. 
By \cref{ex:1.4.1} we see that this third zero $z_3$ must be such that $z_1 + z_2 + z_3 + \Lambda = 0 + \Lambda$. 
Hence $z_3 = - z_1 - z_2$. Its corresponding point on the elliptic curve is 
$(\wp_\Lambda(z_3), \wp'_\Lambda(z_3)$. Then so $P + Q = (\wp_\Lambda(z_3), - \wp'_\Lambda(z_3)) = (\wp_\Lambda(-z_3), \wp'_\Lambda(-z_3)$. 
So the map respects the group action.

From now on we will often say complex elliptic curve, instead of complex torus to highlight their interpretation as elliptic curves.

\todo{Optionally write about maps between complex curves}.
